\Title

### 2013-01-22 synmut.tex
new figures from simulations

The main result of the simulations is that genetic draft can indeed bring weakly
deleterious mutations to high frequencies and result in a dependence of the
fixation probability on initial frequency that is compatible with observations.
First of all, since neutral mutations are much more likely to rise to high
frequency than deleterious ones, the majority of the synonymous mutations needs
to be slightly deleterious observe a significant reduction of $P_\text{fix}$.
In order to further quantify the reduction in fixation probability, we look at
the difference between the neutral curve ($P_\text{fix}(\nu) = \nu$) and the
measured fixation probability and calculate its area (see inset of
\FIG{simfixpvar}). The minimal and maximal values for this area are zero
(neutral-like curve) and 0.5 (no fixation at all), respectively. The HIV data
correspond to an area under the diagonal of approximately 0.2 for synonymous
changes, and a very small area over the diagonal for nonsynonymous changes.
Various simulation curves are shown in \FIG{simfixpvar}. Then, in
\FIGS{simheat1}{simheat2}, we explore the parameter space: the combinations that
yield areas close to the experimental result are roughly indicated by ellipses.
The two crucial parameters that control the fixation probability are the
following: (a) the deleterious effects of hitchhikers compared to the beneficial
effects of escape mutants, and (b) the density of escape mutations. Intuitively,
a higher density of escape mutations (i.e., epitopes) enables a larger degree of
genetic draft, because escape mutations from different epitopes start to combine
and their effects add up. In \FIG{simheat1}, we show that this is indeed the
case in simulations.

\begin{figure}
\begin{center}
\subfloat{\includegraphics[width=0.9\linewidth]{fixation_loss_shortgenome_distance_ada_frac_del_eff_coi_various.pdf}
\label{fig:simfixpvar}}\\
\subfloat{\includegraphics[width=0.9\linewidth]{fixation_loss_shortgenome_area_ada_frac_del_eff_coi_0_01_nescepi_6_heat.pdf}
\label{fig:simheat1}}\\
\subfloat{\includegraphics[width=0.9\linewidth]{fixation_loss_shortgenome_area_ada_frac_del_eff_coi_0_01_nescepi_6_nonsyn_heat.pdf}
\label{fig:simheat2}}
\caption{\comment{USE A GOOD FIG WITH THE CORRECT $s_a$!} The depression in $P_\text{fix}$ depends on the deleterious effect size
 of the synonymous alleles (panel A). Simulations on the escape competition
 scenario show that the density of selective sweeps and the size of the
 deleterious effects of synonymous mutations are the main driving forces of the
 phenomenon. A convex fixation probability is recovered, as seen in the data,
 along the diagonal (panel B): more dense sweeps can support more deleterious
 linked mutations. The density of sweeps is limited, however, by the
 nonsynonymous fixation probability, which is quite close to neutrality (panel
 C). Moreover, strong competition between escape mutants is required, so that
 several escape mutants are ``found'' by HIV within a few months of antibody
production.}
\label{fig:simheat}
\end{center}
\end{figure}

>>>>

The main result of the simulations is that genetic draft can indeed bring weakly
deleterious mutations to high frequencies and result in a dependence of the
fixation probability on initial frequency that is compatible with observations.
First of all, since neutral mutations are much more likely to rise to high
frequency than deleterious ones, the majority of the synonymous mutations needs
to be slightly deleterious observe a significant reduction of $P_\text{fix}$.
In order to further quantify the reduction in fixation probability, we look at
the difference between the neutral curve ($P_\text{fix}(\nu) = \nu$) and the
measured fixation probability and calculate its area (see inset of
\FIG{simfixpvar}). The minimal and maximal values for this area are zero
(neutral-like curve) and 0.5 (no fixation at all), respectively. The HIV data
correspond to an area under the diagonal of approximately 0.2 for synonymous
changes, and a very small area over the diagonal for nonsynonymous changes.
Various simulation curves are shown in \FIG{simfixpvar}. Then, in
\FIG{simsfig}, we explore the parameter space: the combinations that
yield areas close to the experimental result are shown.
The two crucial parameters that control the fixation probability are the
following: (a) the deleterious effects of hitchhikers compared to the beneficial
effects of escape mutants, and (b) the density of escape mutations. Intuitively,
a higher density of escape mutations (i.e., epitopes) enables a larger degree of
genetic draft, because escape mutations from different epitopes start to combine
and their effects add up. In \FIG{simheat1}, we show that this is indeed the
case in simulations.

\begin{figure}
\begin{center}
\subfloat{\includegraphics[width=\linewidth]{simulations_graduallydel.pdf}
\label{fig:simfixpvar}}\\
\subfloat{\includegraphics[width=\linewidth]{simulations_syn.pdf}
\label{fig:simsfig}}
\caption{(Panel A) The depression in $P_\text{fix}$ depends on the deleterious
effect size of the synonymous alleles. This parameter also reduces synonymous
diversity (first inset). The depression found in HIV data is about 0.2.
(Panel B) To assess the parameter space that affects synonymous fixation and
diversity, we run many simulations with random parameters for deleterious effect
size, fraction of deleterious synonymous sites, average size of escape
mutations (color, blue ($10^{-2.5}$) to red ($10^{-1.5}$)), and rate of
introduction of new epitopes (marker size, from $10^{-3}$ to $10^{-2}$ per
generation). Mostly simulations with small deleterious effects and relatively
high deleterious fractions of synonymous mutations reproduce synonymous HIV-like
fixation probability and diversity.}
\label{fig:simheat}
\end{center}
\end{figure}
